### 2018-01-02 reply_to_reviewer3.tex

\documentclass[]{article}

\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage[symbol]{footmisc}
\newcommand{\red}[1]{\textbf{\color{red} #1}}

\usepackage[margin=1.5in]{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5\baselineskip}

%opening
\title{Response to Reviewer 3}
\author{}
\date{}

\begin{document}

\maketitle

We are pleased with the positive assessment of our work by the reviewer and thank him/her for providing constructive input. The responses to the reviewer's comments and adaptations in the revised manuscript are discussed below.

\hrulefill

\paragraph{Reviewer} \textit{This paper presents an interesting study on yaw and induction control of wind farms. The analysis is reliable, and the paper is well written.}

\section*{Comments}
\hrulefill

\paragraph{1. Reviewer} \textit{Page 11, line 325, “additional control cases Ystat and Ymndr…” Although explanations of Ystat and Ymndr can be found later, e.g. on page 20, line 521. It is suggested that these control cases are explained upon their first appearance. In the caption of Figure 6, the case names should be noted as well.}

\paragraph{Response}

\hrulefill

\paragraph{2. Reviewer} \textit{Some figures are difficult to inspect closely. For example, Figures 7, 11, and 13 contain color lines, and the sizes are too small. The authors can present the results in a better manner.}

\paragraph{Response}

\hrulefill

\paragraph{3. Reviewer} \textit{In Figures 9 and 15, the unit of PSD should be added.}

\paragraph{Response}

\hrulefill

\paragraph{4. Reviewer} \textit{Table 2 gives a summary of the effect different control strategies. Why does the meandering yaw case have reduced power gains in the turbulent condition? How do the authors judge the uncertainties in their analysis? For example, by extending the optimization time window or by refining the cell size, will the conclusion be different?}

\paragraph{Response}

\hrulefill

\end{document}